### 2018-07-03 pack/prv/opt/tokipona/aux/article/acm/samplebody-journals.tex
revisada parte de sh

% \clearpage
\subsection{Syntax highlighting}\label{shigh}
The same package~\cite{tokipona}
has a Vim~\cite{vim} syntax highlighting plugin
for Toki Pona,
routines to arrange the coloring schemes
(i.e. to yield the text files with the specifications for which tokens
to relate to which coloring groups),
and instructions for installing and using
the resulting syntax highlighting for Toki Pona texts~\cite{tokipona}.
An online syntax highlighting gadget is found at~\cite{tpNetSH},
and the solution here described presents a number of enhancements:
it is capable of coloring all morphosyntactic classes;
it might be fine-tuned in the colors and their relations to sets of
tokens; it is designed to be used within Vim
and might be exported as HTML; the highlighting scheme is promptly
rendered by a script; the resulting script might be used in combination
with virtually any color scheme.

Basically, the resulting syntax highlighting
distinguishes the words among the morphosyntactic
classes according to the official dictionary as given in
Table~\ref{foobar}.
A word often belongs to more than one class,
thus precedence might be set by the user.
Also, some classes might be further refined (e.g. words beginning with
vowels) or joined (e.g. distinguishing only particles and the rest,
or particles and prepositions and the rest).
The colors are also promptly changed according to~\cite{vim}
and exemplified in the package documentation.

Currently, the Python package synthesizes the
syntax file through the\\ \tttt{makeVimSyntax.py} script.
The user has control of class precedence and
merging and further details through tweaking such routine.
The choice of precise coloring schemes
involves fine tuning the color scheme being
used in Vim (such as 'blue', 'elflord' and 'gruvbox'),
and Vim's highlighting schemes as described in~\cite{vim}.
In summary,
the usage of the package and plugin might be performed
through the following actions:
\begin{itemize}
  \item Installation of the plugin.
  \item Tweaking of the syntax file by hand.
  \item Running the \tttt{makeVimSyntax.py}
    Python script to generate a new \tttt{tokipona.vim} syntax file
    according to other settings.
  \item Write a file with the \tttt{.tokipona} or \tttt{.tp} extension
    (inside Vim) using Toki Pona words.
    Reload the highlighting scheme using \tttt{:e} whenever you
    change the syntax file by hand or through the Python script.
  \item Access the used highlighted groups with \tttt{:syntax}.
    Access all the highlighting groups with \tttt{:so
    \$VIMRUNTIME/syntax/hitest.vim} or \tttt{:hi}.
    Change the coloring of a set of terms by associating
    a used group (e.g. tokiponaADJECTIVE) to an existing group (e.g.
    Visual) such as in \tttt{:highlight link tokiponaADJECTIVE Visual}.
\end{itemize}

The syntax file has an association of TP
words with highlighting groups.
It also holds associations of these groups to other groups
to use their color settings.
Therefore, one is able to use various color schemes
in syntax-highlighted TP texts, such as in Figure~\ref{fig:syn}.
A Vim user might run e.g. \tttt{:colorscheme blue},
\tttt{:colorscheme solarize}, \tttt{:colorscheme gruvbox} or
\tttt{:colorscheme elflord} to see the same text colored
with different color schemes (association of colors to sets of tokens).
To interfere directly on the colors chosen,
\tttt{:highlight Normal guifg=\#00000 guibg=\#0000ff}
will change the standard foreground (text) color to pure black
and background to pure blue.
See \tttt{:h gui-colors} and \tttt{:h highlight}
for the way you might edit colors directly.
Useful commands are given as commented lines
in the end of the syntax file. % TTM
One might obtain an HTML file with the colored TP text
using the \tttt{:TOhtml} command.

>>>>

% \clearpage
\subsection{Syntax highlighting}\label{shigh}
The same package~\cite{tokipona}
has a Vim~\cite{vim} syntax highlighting plugin
for Toki Pona.
The Python scripts hold
routines to arrange the coloring schemes
(i.e. to specify which tokens
are related to which coloring groups)
which are stored in the \tttt{syntax/tokipona.vim} file.
An online syntax highlighting gadget is found at~\cite{tpNetSH},
and the solution here described presents a number of enhancements:
it is capable of coloring all morphosyntactic classes;
it might be fine-tuned in the colors and their relations to sets of
tokens; it is designed to be used within Vim
and might be exported as HTML; the highlighting scheme is promptly
rendered by a script; the resulting script might be used in combination
with virtually any color scheme.

Basically, the resulting syntax highlighting
distinguishes the words among the morphosyntactic
classes according to the official dictionary as given in
Table~\ref{foobar}.
Also, some classes might be further defined (e.g. words beginning with
vowels) or joined (e.g. distinguishing only particles against rest,
or particles and prepositions against the rest).
The colors are also promptly changed according to~\cite{vim}
and exemplified in the package documentation,
specially the \tttt{tokipona/syntax/tokipona.vim} syntax file

Currently, the Python package synthesizes the
syntax file through the\\ \tttt{syntax.TPSynHigh} class.
The user has control of class precedence,
merging and further details through tweaking such routine.
The choice of precise coloring schemes
involves fine tuning the color scheme being
used in Vim (such as 'blue', 'elflord' and 'gruvbox'),
and Vim's highlighting schemes as described in~\cite{vim}.
The usage of the package and plugin might be performed,
in summary, through the following actions:
\begin{itemize}
  \item Installation of the plugin, so that
    syntax highlighting of TP texts will be performed in any \tttt{*.tp} or \tttt{*.tokipona} file.
  \item Tweaking of the syntax file by hand. Or
  \item Instantiating the \tttt{syntax.TPSynHigh} class and then executing the \tttt{mkSynHighFile()} method
    of the resulting object.
    This generates a new \tttt{tokipona.vim} syntax file, at the right location in Vim folders,
    according to object settings.
  \item Write a file with the \tttt{.tokipona} or \tttt{.tp} extension
    (inside Vim) using Toki Pona words.
    Reload the highlighting scheme using \tttt{:e} whenever you
    change the syntax file by hand or through the Python routines.
  \item Access the used highlighted groups with \tttt{:syntax}.
    Access all the highlighting groups with \tttt{:so \$VIMRUNTIME/syntax/hitest.vim} or \tttt{:hi}.
    Change the coloring of a set of terms by associating
    a used group (e.g. \tttt{tpADJECTIVE}) to an existing group (e.g.
    Visual) such as in \tttt{:highlight link tpADJECTIVE Visual}.
    This may also be achieved changing \tttt{syntax.TPSynHigh().colors}
    attribute and then executing \tttt{mkSynHighFile()}.
\end{itemize}

\noindent Such procedures are exemplified in \tttt{basic/syntaxhighlighting.py}.

The syntax file has an association of TP
tokens with highlighting groups (HIGs).
It also holds associations of these HIGs to other HIGs 
to use their color settings.
Thus, one may change: associations of tokens to HIGs;
associations of regex patterns to HIGs;
associations of HIGs to other HIGs;
whole color schemes or just sections of it.
The last of these tasks is promptly performed using 
the \tttt{space-c-s} and \tttt{space-c-c} normal Vim commands
supplied by the color plugin of the PRV framework~\cite{vim}.
The other tasks are more easily performed by
scripting Vim, Python import and package source code,
depending on the user ability and intended tasks.

The use of different color schemes within the same syntax
file is exemplified in Figure~\ref{fig:syn}.
By copying only the \tttt{syntax/tokipona.vim} and \tttt{ftdetect/tokipona.vim} files,
a user might benefit from the standard Vim commands.
E.g. \tttt{:colorscheme blue},
\tttt{:colorscheme solarize}, \tttt{:colorscheme gruvbox} or
\tttt{:colorscheme elflord} to see the same text colored
with different color schemes (association of colors to sets of tokens).
To interfere directly on the colors chosen,
\tttt{:highlight Normal guifg=\#00000 guibg=\#0000ff}
will change the standard foreground (text) color to pure black
and background to pure blue.
See \tttt{:h gui-colors} and \tttt{:h highlight}
for the way you might edit colors directly.
One might obtain an HTML file with the colored TP text
using the \tttt{:TOhtml} command.
